---
title: "How to Use Geocoding Functions"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE,echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)

library(tidyr)

library(tidyselect)
```

```{r}
library(DT)
```

## Introduction

This geocoding function takes in address data and returns Area Deprivation Index, Child Opportunity Index, and Social Vulnerability Index values. The geocoding function `geo_2020()` uses 2020 US Census boundaries. The function works by utilizing the censusxy package which allows for batch geocoding through the Census geocoder without an API key.

## Dataset Rules

Address data should be inputted with the address separated into 'street', 'city', 'state', and 'zip' columns. Make sure the column names match the names used in this dataset. Addresses should be stripped of apartment numbers if applicable and any NA values in the data set should be removed. Additionally zip codes should be entered as numbers rather than characters.

```{r}
library(tibble)

x <- tibble(street = c("1501 Washington Ave", "875 N Michigan Ave", 
                       "1901 W Madison St", "8327 N Galena Rd", "951 Chicago Ave"), 
            city = c("Cairo", "Chicago", "Chicago", "Peoria", "Oak Park"),
            state = c("IL", "IL", "IL", "IL", "IL"),
            zip = c(62914, 60611, 60612, 61615, 60302))

datatable(x)
```

For this example we took five addresses formatted correctly and inputted them into the function.

## Getting the Function

To get `geo_2020()` in your R environment run this code sourcing from the Github repository [pltu06/geocoding](https://github.com/pltu06/geocoding/blob/main/src/geo_2020.R).

```{r}
source("https://raw.githubusercontent.com/pltu06/geocoding/main/src/geo_2020.R")
```

## Running the Function

The `download` argument determines whether or not you need to download the index data. If you do not have the index data on your computer you should set `download = TRUE`. `geo_2020()` has one year selector argument. `coi_year` determines which year of COI data you would like to access. The function can output data for any year from 2012 to 2023.

::: callout-warning
The download can stop if you have poor internet connection. Otherwise it should take roughly 20-30 seconds.
:::

```{r}
data_2020 <- geo_2020(data = x, download = TRUE, coi_year = 2023)
```

## Data Output

`geo_2020()` outputs thirteen variables along with the original address data. It outputs twenty-two index variables: COI national, COI domains national (of which there are three), COI subdomains national (of which there are fourteen), ADI state, ADI national, SVI on the tract level, and SVI on the block group level. The COI values are z-scored, the SVI values are percent rank, and the ADI values are percentiles. It also outputs five geographic variables: longitude, latitude, tract number, tract FIPS code, and block group FIPS code.

```{r, echo = FALSE}
data_2020 <- data_2020%>%
  mutate(across(.cols = all_of(c("svi_block", "z_COI_nat", "z_ED_nat", "z_HE_nat", "z_SE_nat", "z_ED_EC_nat", "z_ED_EL_nat", "z_ED_ER_nat", "z_ED_SP_nat", "z_HE_EP_nat", "z_HE_HR_nat", "z_HE_SE_nat", "z_HE_HE_nat", "z_SE_EI_nat", "z_SE_EO_nat", "z_SE_ER_nat", "z_SE_HQ_nat", "z_SE_SR_nat", "z_SE_WL_nat")), .fns = ~round(.x, digits = 4)))

datatable(data_2020)
```
